% Template for a Thesis
%
% Acknowledgements.tex
%
% Acknowledgements

\begin{center}
    \huge
    Acknowledgements
    \vspace{1cm}
\end{center}

The development of this thesis was hit by the COVID-19 pandemic, reclusing me in my hometown and depriving me from face to face contact with my supervisor and exchange of ideas with other students. Writing this thesis has been a more solitary endeavor than what I would have hoped, but it still was grounded in the support of many people.

First of all, my biggest appreciation goes to my supervisor Masoud, with whom I had weekly video chats to discuss the progress, brainstorm ideas and discuss next steps. He has always been very helpful and attentive, and especially empathetic with my confinement when he displayed colorful and paradisiac beaches as his background, to make me jealous. Despite his many other obligations he has found time for me, and provided guidance, direction or resources. It has been a pleasure to work with you.

On the other side, my family has been always been a pillar for me, and more so during these difficult times. They have provided distraction, entertainment, cheering up, sharing ideas, progress or lack thereof, and helping me through the so common situation when a bug torments you for days or weeks, appearing in your nightmares, making you obsessed until people around you start worrying if you are having a breakdown. They ask you to explain it, and talking about it to someone outside of circle of competence makes you explain it in an alternative way, and they ask questions that make you think about the problem in a completely different way, and at some point you think: maybe I could quickly try this, wait a second...

Last but not least, friends and loved ones have also provided great support. Thank you for the regular checkups, early draft reviews, ranting opportunities, the burgers in the park. You have given me a hand to hold, things to do, a room to sit in surrounded by my favourite view.

% end of file